Supporting Information to Estimating the Initial Susceptible Fraction in Dengue Dynamics"

### 2014-11-04 S1_lm1.tex
changes to S1, long forgotten, sorrey :-(

\subsection*{A remark on prior distributions and tail behavior of the 
distribution of $R_t$}
\label{sec:tails}

In order to decide which approach to take, it may be of use analysing the variance and tail behaviour of the derived distributions for \rr. 
Consider the case of using a flat $Uniform(0, 1)$ prior for $\theta_t$.
With $a_0 = b_0 = 1$, $a_1 = a_2$ and $b_1 = b_2 + 1$.
The beta prime (inverse beta distribution) will have heavier tails compared to the conditional distribution in (\ref{eq:densityNishiura}), thus providing more conservative confidence/credibility intervals.  
As a side note, the Bayesian approach presented in Section~\ref{sec:mantel} will give similar results to those of \citet{wilson} and \citet{wilson} for $Y_{t+1}$ and $Y_t >> 1$.
Under the flat  uniform prior for $\theta_t$, the Bayesian posterior credibility interval is nearly indistinguishable from the confidence interval proposed by \citet{clopper} for $Y_{t+1}, Y_t > 20$.
Note that the $Beta(1, 1)$ uniform prior for $\theta_t$ constitutes a poor prior choice mainly because the induced distribution for \rr is only well-defined for $b_0 > 2$.

The main advange of the Bayesian approach is that one can devise prior distributions for $\theta_t$ taking advantage of the intuitive parametrisation and flexibility of the beta family of distributions.
Prior elicitation can also be done for \rr and the hyperparameters directly plugged into the prior for $\theta_t$. 
One can, for example, choose a priori mean and variance for \rr and find $a_0$ and $b_0$ that satisfy those conditions.
Let $m_0$ and $v_0$ be the prior expectation and variance for $R_t$. 
After some tedious algebra one finds
\begin{align}
\label{eq:elicitation}
a_0 &= \frac{m_0v_0 + m_0^3 + m_0^2}{v_0} \\
b_0 &= \frac{2v_0 + m_0^2 + m_0}{v_0}
\end{align}
If one wants only to specify $m_0$ and a coefficient of variation $c$ \footnote{$c = \sqrt{v_0}/ m_0$.} for $R_t$ \textit{a priori}, some less boring algebra gives:
\begin{align}
\label{eq:elicitationcv}
a_0 &= \frac{m_0^3c^2 + m_0^3 + m_0^2}{m_0^2c^2} \\
b_0 &= \frac{2m_0^2c^2 + m^2 + m}{m_0^2c^2}
\end{align}
\newpage
\bibliography{lm1}
\end{document}

>>>>

\subsection*{A remark on prior distributions and tail behavior of the 
distribution of $R_t$}
\label{sec:tails}
There are a number of approaches to deriving the distribution of \rr.
One could for example  use the distribution in equation A6 of~\citet{nishiura} 
or estimate \rr using the approach described in the main text~\citep{mantel}.
In order to decide which approach to take, it may be of use analyzing the 
variance and tail behavior of the derived distributions for \rr. 
Consider the case of using a flat $Uniform(0, 1)$ prior for $\theta_t$.
With $a_0 = b_0 = 1$, $a_1 = a_2$ and $b_1 = b_2 + 1$.
The beta prime (inverse beta distribution) will have heavier tails compared to 
the conditional distribution proposed by~\citet{nishiura}, thus providing more 
conservative confidence/credibility intervals.  
As a side note, the Bayesian approach presented in this 
paper will give similar results to those of 
\citet{wilson} and \citet{wilson} for $Y_{t+1}$ and $Y_t >> 1$.
Under the flat  uniform prior for $\theta_t$, the Bayesian posterior credibility interval is nearly indistinguishable from the confidence interval proposed by \citet{clopper} for $Y_{t+1}, Y_t > 20$.
Note that the $Beta(1, 1)$ uniform prior for $\theta_t$ constitutes a poor prior choice mainly because the induced distribution for \rr is only well-defined for $b_0 > 2$.

An advantage of the Bayesian approach is that one can devise prior 
distributions for $\theta_t$ taking advantage of the intuitive parameterization 
and flexibility of the beta family of distributions.
Prior elicitation can also be done for \rr and the hyperparameters directly plugged into the prior for $\theta_t$. 
One can, for example, choose a priori mean and variance for \rr and find $a_0$ and $b_0$ that satisfy those conditions.
Let $m_0$ and $v_0$ be the prior expectation and variance for $R_t$. 
After some tedious algebra one finds
\begin{align}
\label{eq:elicitation}
a_0 &= \frac{m_0v_0 + m_0^3 + m_0^2}{v_0} \\
b_0 &= \frac{2v_0 + m_0^2 + m_0}{v_0}
\end{align}
If one wants only to specify $m_0$ and a coefficient of variation $c$ \footnote{$c = \sqrt{v_0}/ m_0$.} for $R_t$ \textit{a priori}, some less boring algebra gives:
\begin{align}
\label{eq:elicitationcv}
a_0 &= \frac{m_0^3c^2 + m_0^3 + m_0^2}{m_0^2c^2} \\
b_0 &= \frac{2m_0^2c^2 + m^2 + m}{m_0^2c^2}
\end{align}

This approach thus makes possible to incorporate epidemiological knowledge 
about disease biology (e.g. the magnitude of $R_0$) into the computation of \rr.
This may prove particularly important when disease counts are low and/or close 
to the detection threshold.
\newpage
\bibliography{lm1}
\end{document}
